 

### 2021-04-26 vertical-thermal-review.Rmd
Update vertical-thermal-review.Rmd

Typical values across latitude and climate are ~45-50$^\circ$C (O' sullivan et al. 2017). Across forest species, in Panama, Columbia and across latitudes, $T_{50}$ and $T_{crit}$ decreased with elevation [@feeley_thermal_2020;  Slot et al 2020; O' sullivan et al. 2017] and varies with leaf traits [Sastry et al. 2018, @zhang_photosynthetic_2012; Slot et al. 2020]. Species with large leaves, and slow heating and cooling time [higher thermal time constant] showed greater $T_{crit}$. On the other hand, leaves of species with greater construction costs, observed through LMA, had greater $T_{50}$ compared to species with low LMA [Slot et al. 2020]. However, higher photosynthetic heat tolerance does not necessary increase leaf thermal safety margins, rather, in across species sun leaves, higher photosynthetic heat tolerance experienced higher maximum temperatures, and showed low optimum temperature for C assimilation [$T_{opt}$] and had narrower thermal safety margins [Perez and Feeley, 2020]

@@

While dark and light respiration tends to increase with light exposure, and therefore along the vertical gradient, with greater respiration in canopy leaves [@bolstad_foliar_1999; Weerasinghe et al.2014], temperature sensitivity of respiration, $Q_{10}$, shows conflicting findings with no definite trend.

>>>>

Typical values across latitude and climate are ~45-50$^\circ$C (O' sullivan et al. 2017). Across forest species, in Panama, Columbia and across latitudes, $T_{50}$ and $T_{crit}$ decreased with elevation [@feeley_thermal_2020;  Slot et al 2020; O' sullivan et al. 2017] and varies with leaf traits [Sastry et al. 2018, @zhang_photosynthetic_2012; Slot et al. 2020]. Species with large leaves, and slow heating and cooling time [higher thermal time constant] showed greater $T_{crit}$. On the other hand, leaves of species with greater construction costs, observed through LMA, had greater $T_{50}$ compared to species with low LMA [Slot et al. 2020]. However, higher photosynthetic heat tolerance does not necessary increase leaf thermal safety margins. Rather, in across species sun leaves, higher photosynthetic heat tolerance experienced higher maximum leaf temperatures, and showed low optimum temperature for C assimilation [$T_{opt}$] and therefore had narrower thermal safety margins [Perez and Feeley, 2020.]

@@

While dark and light respiration tends to increase along the vertical gradient, with greater respiration in canopy leaves [@bolstad_foliar_1999; Weerasinghe et al.2014], temperature sensitivity of respiration, $Q_{10}$, shows conflicting findings with no definite trend.
